
\hrule
\vspace{1em}

Requirements: 

\textit{``defines the plan formulated by the student and the committee regarding (at least)}:

\begin{itemize}
\item \textit{the topic and scope of the dissertation,}
\item \textit{method and scope of data collection, and}
\item \textit{analysis strategies.}
\end{itemize}

\vspace{1em}
\hrule
\vspace{2em}

I see myself as a transdisciplinary researcher: sprawled across worlds, but following a common thread. The closest I've come to a simple description of my focus is \textit{communicative systems}: how agents or components interact to create systems of meaning that, reciprocally, structure their interaction within and across what appear as emergent layers. I also believe that the practice of science itself is deeply political, reflective of the systems of belief and power interacting across interlocking institutions and organizations, and have increasingly oriented my work towards understanding and directly counteracting the ways that commercial capture of science poison every part of our work.

This dissertation will be in three chapters, roughly tracing the development of my thoughts over the past seven years.

\textbf{Chapter 1: \hyperref[sec:phonemes]{Phonetic Perception}} - My work here starts and ends with studying a mouse model of auditory phonetic perception, comprising behavioral neuroscience, an attempted merger of several strains of linguistic, philosophical, and neuroscientific thought, and in the final few months a touch of neurophysiology. 

\textbf{Chapter 2: \hyperref[sec:autopilot]{Autopilot}} - That led, perhaps unwisely, to the development of Autopilot: a Python framework for behavioral neuroscientific experiments based in distributed computing, but in learning lessons from history and critical technology scholars has been mutating into something more like a technical knowledge organization system. 

\textbf{Chapter 3: \hyperref[sec:infrastructure]{Infrastructure}} - Finally in my last year my work has focused on the broader infrastructure and practice of science, and weaving between many fields I have tried to plot a course away from the subordination of basic science to surveillance and information capitalism. 

These three chapters reflect my study of three interconnected emergent layers of scientific practice: science \textit{qua} science, systems that enable science, and metasystems that structure the systems of science. 

\clearpage

\input{_speech}

\input{_autopilot}

\input{_infrastructure}

\input{_timeline}

\clearpage

\input{_resources}
